Notes on Wide Gap

### 2016-04-10 paper/wide_gap_notes.tex
fixed G matrix
controlled gap width nondimensionalization -- see question

The momentum equation must be nondimensionalized. Here are the definitions of all of the dimensional components:

@@

\beq
\mathcal{G} = \left[\begin{matrix}
0 & 3\Omega(r) & 0 & 0 \\
-\frac{4}{r} \Omega(r)  - \partial_r \Omega(r)  & 0 & 0 & 0 \\
0 & 0 & 0 & 0 \\
0 & 0 & \frac{2}{r} \Omega(r)  + \partial_r \Omega(r)  & 0
\end{matrix}\right]
\eeq

>>>>

\subsection{Controlled gap width nondimensionalization}

The idea is to parameterize the gap width rather than simply nondimensionalizing lengths by a point in the center of the channel $r_0$, as we did below.

There are many ways to do this but one option is to define

\beq
\widetilde{r} = r_0 + \widetilde{x}
\eeq

Where tildes indicate dimensional quantities. Then for $\widetilde{x}$ to stay Cartesian, we need

\beq
\widetilde{r} - r_0 \ll r_2 - r_1 = d
\eeq

where $r_2 = r_1 = d$ is the gap width. Also define $\widetilde{x} \ll d$ and $d/r_0 << 1$ for the small gap approximation. Thus when we nondimensionalize,

\beq
r = \frac{\widetilde{r}}{r_0} = 1 + \frac{\widetilde{x}}{r_0}
\eeq

Note that we could control the gap width in terms of a small parameter $\epsilon x$ instead of a fraction $\frac{x}{r_0}$, but I believe both should work the same way.

In the example Jeff worked out before, we looked at the base velocity profile in the thin gap limit. The rotation profile is given by $\Omega(r) = \Omega_0 r^{-q}$ in dimensionless quantities, and the background flow $u = r\Omega(r)$, so in our new notation:

\beq
u = \left(1 + \frac{\widetilde{x}}{r_0}\right) \Omega_0 \left(1 + \frac{\widetilde{x}}{r_0}\right)^{-q} = \Omega_0 \left(1 + \frac{\widetilde{x}}{r_0}\right)^{1-q}  
\eeq

Taylor expanding the exp(1-q) quantity, we arrive at

\beq
u = \Omega_0 \left(1 + \frac{\widetilde{x}}{r_0}\right) - \Omega_0 q \frac{\widetilde{x}}{r_0}
\eeq

Then if we define this with respect to a rest frame $\Omega_0 \left(1 + \frac{\widetilde{x}}{r_0}\right)$ and define a dimensionless $x = \widetilde{x}/r_0$, we get the Umurhan+ base flow $u = - \Omega_0 q x$.

This makes sense to me but I'm having trouble extending it to the actual equations. This implies, for example, that the term $\frac{1}{\eta} \nabla^2 B$, which in the $\hat{\phi}$ equation is

\beq
\left[\nabla^2 B\right]_\phi = \nabla^2 B_\phi - \frac{B_\phi}{r^2},
\eeq

should become

\beq
\nabla^2 B_\phi - B_\phi \left(1 - \frac{2 \widetilde{x}}{r_0}\right)
\eeq

after substitution and Taylor expansion of the $r^{-2}$. But clearly there's no $- B_\phi \left(1 - \frac{2 \widetilde{x}}{r_0}\right)$ in the thin gap equation. What am I missing?? Is this not the right place to be substituting in the new definition of r??

\subsection{Typical nondimensionalization}
The momentum equation must be nondimensionalized. Here are the definitions of all of the dimensional components:

@@

\beq
\mathcal{G} = \left[\begin{matrix}
0 & 0 & \frac{2}{\beta} \frac{1}{r} \left(\partial_r^2 - \frac{1}{r}\partial_r\right) & 0 \\
0 & 0 & 0 & \frac{2}{\beta}\\
1 & 0 & 0 & 0\\
0 & 1 & 0 & 0
\end{matrix}\right] 
\eeq
